Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-04 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
hs: emphasized diff between us and carullo

Following the ParSpec expansion, our QNM frequencies are expressed through functions,
%
\begin{subequations}
\begin{align}
\omega_k &= \omega_k(m_1, m_2, \chi_1, \chi_2,\ell, \{\delta \omega_k^{(j)}\}),\\
\tau_k   &= \tau _k(m_1, m_2, \chi_1, \chi_2, \ell, \{\delta \tau_k^{(j)}\}),
\end{align}
\end{subequations}
%
where we fixed a certain value of $p$, and the estimates of $(m_1,\, m_2,\, \chi_1,\, \chi_1)$ are used to predict
the final BH mass $M_f$ and spin $\chi_f$ through~\cite{Taracchini:2013rva,Hofmann:2016yih}.
%
Additionally, both $\omega_k$ and $\tau_k$ depend on the coupling constant $\ell$
defined in Eq.~\eqref{eq:def_gamma} as well as the sets, $\{\delta
\omega_k^{(n)}\},$ and $\{\delta \tau_k^{(n)}\}$, where $n$ ranges from zero to
the power of spin in Eqs.~\eqref{eq:kerr_expansion} up to which we include
corrections.
%
For example, if we restrict ourselves to corrections with $n=0$, or
$\chi_f^0$, to the $k=(2,2,0)$ mode only, then we would just have two
additional parameters, $\{\delta \omega_0^{(0)}, \delta \tau_0^{(0)}\}$.
%
In this paper, we will restrict ourselves to corrections to the $n=0$ and
$n=1$ terms, i.e., leading and next-to-leading order corrections to the spin,
$\chi_f$.

Using this parameterized waveform model, which we will call $\pSEOB$ in this
paper, we infer bounds on our beyond-GR parameters for specific cases of
modified gravity theories presented Sec.~\ref{sec:review_theories}.
%
We detail these results in Sec.~\ref{sec:results}.
%
\hs{Should we call this waveform differently to distinguish it from your previous papers?}
%
%In this paper, we will restrict ourselves to the $k=(2,2,0)$ QNM and corrections to the $j=0$ and $j=1$ terms, i.e., leading and next-to-leading order corrections to the spin, $\chi_f$. Assuming a prior which is uniform on the beyond-GR parameters, $(\ell, \{\delta \omega_k^{(j)}\},\{\delta \tau_k^{(j)}\})$, we employ a Bayesian framework to stochastically sample over the parameter space using the \texttt{LALInference} algorithm.~\cite{lallsuite}.
%
%The bounds on these beyond-GR parameters, for specific cases of modified gravity theories outlined in Sec.~\ref{sec:review_theories}, are detailed in Sec.~\ref{sec:results}.

@@

Let us now establish the connection between theory-agnostic framework of the
$\pSEOB$ waveform model and the theory-specific QNM calculations.
%
We begin by rewriting $\omega_{k}$ and $\tau_{k}$, given by Eqs.~\eqref{eq:kerr_expansion} as,
%
\begin{subequations}
\label{eq:qnm_nongr_iso}
\begin{align}
M \omega_{k} &= \gamma \left( \delta\omega^{(0)} \omega^{(0)} + \chi \delta\omega^{(1)} \omega^{(1)} \right)
+ \sum_{n=0}^{N_{\rm max}} \chi^{j} \omega^{j}\,,
\\
\tau_{k}/M   &= \gamma \left( \delta\tau^{(0)} \tau^{(0)} + \chi \delta\tau^{(1)} \tau^{(1)} \right)
+ \sum_{n=0}^{N_{\rm max}} \chi^{(n)} \tau^{(n)}\,,
\end{align}
\end{subequations}
%
We will work exclusively with the $\ell = m = 2$ and $n = 0$ (fundamental mode)
QNM, i.e., $k = \{2,\, 2,\, 0\}$. For this reason we will omit the subcripts
$k$ hereafter and no ambiguity in notation between the overtone index and the
subscript $n$ in the series should arise.
%
We have also pulled out from the sum all beyond-GR corrections, restricting
ourselves to the nonspinning ($n=0$) and linear-order in spin ($n=1$)
corrections to the Kerr QNMs.
%
% These corrections are encoded in the coefficients $\delta\omega^{(n)}$ and $\delta\tau^{(n)}$.

@@

This means that using the $\pSEOB$ waveform model with theory-specific informed
QNMs we reduce the parameter space of beyond-GR parameters to $\ell$ alone.
%
We emphasize that our procedure is different from that of
Ref.~\cite{Carullo:2021dui} which, for a given value of $p$, varied all $\ell$,
$\delta\omega^{(n)}$ and $\delta\tau^{(n)}$ parameters.

@@

We now turn our attention to dCS gravity.
%
In Fig.~\ref{fig:dCS_exec_sum} we show the marginalized posterior distribution
function of $\ell_{\rm CS}$ for both events GW150914 (top panel) and GW200129
(middle panel), with (purple line) and without (black line) spin corrections to
dominant $(2,2,0)$~QNM.
%
Perhaps the most salient feature of these posteriors is that they are not
peaked at $\ell_{\rm CS}$, which may mislead us to believe that we are seeing a
deviation from GR. This is not the case and the situation is that same as the
one we discussed in the previous section for EdGB gravity.
%
We verified that
%
\begin{equation}
    \gamma_{\rm CS} = \left( \frac{c^2 \ell_{\rm CS}}{G M_{\rm s}} \right)^{4}\,,
\end{equation}
%
does indeed peak at zero indicating consistency with GR, similarly to what is shown in Fig.~\ref{} for $\gamma_{\rm GB}$.
%
We also see that in both cases, that the inclusion leading-order-in-spin
correction to the mode, displaces the posteriors towards smaller values of
$\ell_{\rm CS}$. This can be seen more evidently by looking at the location of
posterior peaks.
%
Finally, in the bottom panel, we show the combined result for both events.

>>>>

Following the ParSpec expansion, our QNM frequencies are expressed through functions,
%
\begin{subequations}
\begin{align}
\omega_k &= \omega_k(m_1, m_2, \chi_1, \chi_2,\ell, \{\delta \omega_k^{(j)}\}),\\
\tau_k   &= \tau _k(m_1, m_2, \chi_1, \chi_2, \ell, \{\delta \tau_k^{(j)}\}),
\end{align}
\end{subequations}
%
where we fixed a certain value of $p$, and the estimates of $(m_1,\, m_2,\, \chi_1,\, \chi_1)$ are used to predict
the final BH mass $M_f$ and spin $\chi_f$ through~\cite{Taracchini:2013rva,Hofmann:2016yih}.
%
Additionally, both $\omega_k$ and $\tau_k$ depend on the coupling constant $\ell$
defined in Eq.~\eqref{eq:def_gamma} as well as the sets, $\{\delta
\omega_k^{(n)}\},$ and $\{\delta \tau_k^{(n)}\}$, where $n$ ranges from zero to
the power of spin in Eqs.~\eqref{eq:kerr_expansion} up to which we include
corrections.
%
For example, if we restrict ourselves to corrections with $n=0$, or
$\chi_f^0$, to the $k=(2,2,0)$ mode only, then we would just have two
additional parameters, $\{\delta \omega_0^{(0)}, \delta \tau_0^{(0)}\}$.
%
In this paper, we will restrict ourselves to corrections to the $n=0$ and
$n=1$ terms, that is, only up to the leading-order in spin $\chi_f$ term.

Using this parameterized waveform model, which we will call $\pSEOB$ in this
paper, we infer bounds on our beyond-GR parameters for specific cases of
modified gravity theories presented Sec.~\ref{sec:review_theories}.
%
We detail these results in Sec.~\ref{sec:results}.
%
\hs{Alessandra, Abhirup, should we call this waveform differently to distinguish it from your previous papers?}
%
%In this paper, we will restrict ourselves to the $k=(2,2,0)$ QNM and corrections to the $j=0$ and $j=1$ terms, i.e., leading and next-to-leading order corrections to the spin, $\chi_f$. Assuming a prior which is uniform on the beyond-GR parameters, $(\ell, \{\delta \omega_k^{(j)}\},\{\delta \tau_k^{(j)}\})$, we employ a Bayesian framework to stochastically sample over the parameter space using the \texttt{LALInference} algorithm.~\cite{lallsuite}.
%
%The bounds on these beyond-GR parameters, for specific cases of modified gravity theories outlined in Sec.~\ref{sec:review_theories}, are detailed in Sec.~\ref{sec:results}.

@@

Let us now establish the connection between theory-agnostic framework of the
$\pSEOB$ waveform model and the theory-specific QNM calculations.
%
We begin by rewriting $\omega_{k}$ and $\tau_{k}$, given by Eqs.~\eqref{eq:kerr_expansion} as,
%
\begin{subequations}
\label{eq:qnm_nongr_iso}
\begin{align}
M \omega_{k} &= \gamma \left( \delta\omega^{(0)} \omega^{(0)} + \chi \delta\omega^{(1)} \omega^{(1)} \right)
+ \sum_{n=0}^{N_{\rm max}} \chi^{n} \omega^{(n)}\,,
\\
\tau_{k}/M   &= \gamma \left( \delta\tau^{(0)} \tau^{(0)} + \chi \delta\tau^{(1)} \tau^{(1)} \right)
+ \sum_{n=0}^{N_{\rm max}} \chi^{n} \tau^{(n)}\,,
\end{align}
\end{subequations}
%
We will work exclusively with the $\ell = m = 2$ and $n = 0$ (fundamental mode)
QNM, i.e., $k = \{2,\, 2,\, 0\}$. For this reason we will omit the subcripts
$k$ hereafter and no ambiguity in notation between the overtone index and the
subscript $n$ in the series should arise.
%
We have also pulled out from the sum all beyond-GR corrections, restricting
ourselves to the nonspinning ($n=0$) and linear-order in spin ($n=1$)
corrections to the Kerr QNMs.
%
% These corrections are encoded in the coefficients $\delta\omega^{(n)}$ and $\delta\tau^{(n)}$.

@@

This means that using the $\pSEOB$ waveform model with theory-specific informed
QNMs we reduce the parameter space of beyond-GR parameters to $\ell$ alone.
%
We emphasize that our procedure is different from that of
Ref.~\cite{Carullo:2021dui} which, for a given value of $p$, varied all $\ell$,
$\delta\omega^{(n)}$, and $\delta\tau^{(n)}$ parameters, and then proceeded
to use the posteriors on $\ell$ to draw conclusions to any theory that predict
corrections at a given $p$.
%
This is a critical different between our study and Ref.~\cite{Carullo:2021dui}.
In particular, as we will see in Sec.~\ref{sec:results}, adding theory-specific
information to the ParSpec expansion \emph{can lead to different interpretation
of the final results, even for different theories which predict modifications
with the exponent $p$.}

@@

We now turn our attention to dCS gravity.
%
In Fig.~\ref{fig:dCS_exec_sum} we show the marginalized posterior distribution
function of $\ell_{\rm CS}$ for both events GW150914 (top panel) and GW200129
(middle panel), with different colors distinguishing cases with and without spin corrections to
dominant $(2,2,0)$~QNM.
%
Perhaps the most salient feature of these posteriors is that they are not
peaked at $\ell_{\rm CS}$, which may mislead us to believe that we are seeing a
deviation from GR. This is not the case and the situation is that same as the
one we discussed in the previous section for EdGB gravity.
%
We verified that
%
\begin{equation}
    \gamma_{\rm CS} = \left( \frac{c^2 \ell_{\rm CS}}{G M_{\rm s}} \right)^{4}\,,
\end{equation}
%
does indeed peak at zero indicating consistency with GR, similarly to what is shown in Fig.~\ref{} for $\gamma_{\rm GB}$.
%
We also see that in both cases, that the inclusion leading-order-in-spin
correction to the mode, displaces the posteriors towards smaller values of
$\ell_{\rm CS}$. This can be seen more evidently by looking at the location of
posterior peaks.
%
Finally, in the bottom panel, we show the combined result for both events.
